---
title: "New York Congressional Elections, 1800"
author: "Mapping Early American Elections Project Team"
output: html_document
---

Below are some sample maps and visualizations of the 1800 New York congressional elections for a presentation at the annual meeting of the Society for Historians of the Early American Republic, July 22, 2016.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load-data, include=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(USAboundaries)
library(leaflet)
library(rgdal)
library(scales)
library(ggplot2)
library(ggrepel)

nnv <- read_tsv("data-raw/nnv-tsv/all-votes.tsv")
join_table <- read_csv("data/ny_county_join_table.csv")

names(nnv) <- names(nnv) %>%
  str_to_lower() %>%
  str_replace_all("\\.", "") %>%
  str_replace_all("\\s", "_")

ny_congressional <- nnv %>%
  filter(office == "U.S. House of Representatives",
         state == "New York",
         !is.na(county))

ny_1800 <- ny_congressional %>% 
  filter(date == "1800",
         is.na(town),
         is.na(city))

# ny_1800 %>% 
#   group_by(id, district, county) %>% 
#   mutate(vote = if_else(is.na(vote), 0L, vote),
#          vote_county = sum(vote, na.rm = TRUE),
#          percentage_county = vote / vote_county,
#          won_county = vote == max(vote, na.rm = TRUE)) %>% 
#   group_by(id) %>% 
#   mutate(vote_district_total = sum(vote)) %>% 
#   group_by(id, name_id) %>% 
#   mutate(vote_candidate_total = sum(vote),
#          vote_candidate_percentage = round(vote_candidate_total / vote_district_total, 3),
#          counties_won = sum(won_county)) %>% 
#   group_by(id) %>% 
#   mutate(won_election = vote_candidate_total == max(vote_candidate_total)) 
  
data_for_map <- ny_1800 %>% 
  group_by(affiliation, county) %>%
  summarize(vote = sum(vote),
            district = district[1]) %>% 
  spread(affiliation, vote, fill = 0) %>% 
  mutate(total_vote = Federalist + Republican,
         federalist_percentage = Federalist / total_vote,
         federalist_percentage =
           ifelse(is.nan(federalist_percentage), NA, federalist_percentage),
         deviation = federalist_percentage - 0.5) %>% 
  left_join(join_table, by = c("county" = "nnv_county")) %>% 
  ungroup() %>% 
  mutate(winner = ifelse(Federalist > Republican,
                         "Federalist", "Republican"))

nhgis_join <- read_csv("data/nhgis_nycounty_join_table.csv")
populations <- read_csv("data-raw/population/nhgis0049_ds2_1800_county.csv")

ny_pop <- populations %>% 
  filter(STATE == "New York") %>% 
  mutate(fwm = AAX003 + AAX004 + AAX005)

data_for_map <- data_for_map %>% 
  left_join(nhgis_join, by = c("county" = "nnv_county")) %>% 
  left_join(ny_pop, by = c("nhgis_county" = "COUNTY")) %>% 
  mutate(total_vote = ifelse(total_vote == 0, NA, total_vote),
         turnout = total_vote / fwm) 
```

### Map of votes for Federalist and Republicans by county

```{r, fig.width=8, fig.height=8}
counties_1800 <- us_counties("1800-07-04", states = "New York", resolution = "high")

counties_1800@data <- counties_1800@data %>% 
  left_join(data_for_map, by = c("name" = "ahcb_county")) %>% 
  mutate(total_vote = ifelse(total_vote == 0, NA, total_vote),
         alpha = rescale(to = c(0.5, 1), total_vote)) 

districts_1800 <- readOGR("data-raw/congressional-districts", "ny-1800-districts", verbose = FALSE)

color_scale <- colorNumeric("PRGn", domain = c(0, 1), na.color = "darkgrey")

leaflet(counties_1800) %>% 
  addPolygons(stroke = TRUE, 
              color = "#000000", 
              opacity = 1, 
              weight = 1,
              smoothFactor = 0.5,
              fillColor = ~color_scale(federalist_percentage),
              fillOpacity = alpha,
              popup = ~paste0("<b>County: </b>", county, "<br>",
                              "Federalist vote: ",
                              prettyNum(Federalist, big.mark = ","),
                              " (", round(federalist_percentage * 100, 1),
                              "%)<br>", "Republican vote: ",
                              prettyNum(Republican, big.mark = ","), 
                              " (", round((1 - federalist_percentage) * 100, 1),
                              "%)<br>")) %>% 
  addPolygons(data = districts_1800,
              stroke = TRUE,
              color = "#000000",
              opacity = 1, 
              fill = FALSE, 
              weight = 3,
              smoothFactor = 0.5) %>%
  addLegend(pal = color_scale,
            values = ~c(federalist_percentage, 1),
            labFormat = labelFormat(suffix = "%",
                                    transform = function(x) 100 * x),
            na.label = "NA",
            title = "Federalist %")
```

### Visualizations of Congressional elections in New York 1800

```{r, fig.width=8, fig.size=8}
ny_1800 %>% 
  group_by(id, name_id) %>% 
  summarize(name = name[1],
            affiliation = affiliation[1],
            vote = sum(vote, na.rm = TRUE)) %>% 
  group_by(id) %>% 
  mutate(vote_total = sum(vote, na.rm = TRUE),
         percentage = vote / vote_total,
         winner = vote == max(vote),
         label = ifelse(winner, name, NA)) %>% 
  mutate(district = str_extract(id, "\\d+") %>% 
           as.numeric() %>% ordered(levels = c("10", "9", "8", "7", "4", "3",
                                               "1"))) %>% 
  arrange(affiliation, desc(vote)) %>% 
  ggplot(aes(x = district,
             y = percentage, fill = affiliation,
             label = label)) + 
  geom_bar(stat = "identity", color = "white") +
  coord_flip() +
  scale_fill_manual(values = c("#7fbf7b", "gray", "#af8dc3"),
                    labels = c("Federalist", "NA", "Republican")) +
  geom_hline(yintercept = 0.5, color = "darkgray") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "Competition between parties and candidates",
       subtitle = "New York congressional elections, 1800",
       x = "District",
       y = NULL,
       fill = "Party") +
  scale_y_continuous(labels = percent)
```


```{r, fig.width=8, fig.height=8}
data_for_map %>% 
  mutate(label = ifelse(Federalist > 250 | Republican > 250,
                        county, NA)) %>% 
ggplot(aes(x = Federalist, y = Republican, color = winner,
           size = turnout, label = label)) + 
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "darkgray") +
  geom_text_repel(size = 3.5, color = "black") +
  scale_size_continuous(label = percent) +
  coord_equal() +
  scale_color_manual(values = c("#7fbf7b", "#af8dc3")) +
  theme_minimal() +
  labs(title = "New York congressional elections, 1800",
       subtitle = "Number of votes per county by party",
       color = "Winning party") +
  theme(legend.position = "bottom")
```


```{r fig.width=8, fig.height=8}
data_for_map %>% 
  ggplot(aes(x = turnout, y = federalist_percentage, size = total_vote,
             color = winner, label = county)) +
  geom_point() + 
  theme_minimal() +
  scale_x_continuous(label = percent) +
  scale_y_continuous(label = percent) +
  scale_color_manual(values = c("#7fbf7b", "#af8dc3")) +
  geom_hline(yintercept = 0.5, color = "darkgray") +
  geom_text_repel(size = 3.5, color = "black") +
  labs(title = "Turnout in New York congressional elections, 1800",
       subtitle = "Votes per free white male over 16 by county",
       y = "Federalist %",
       x = "Turnout %",
       color = "Winning party",
       size = "Total votes cast")
  
```

### Map of turnout in New York congressional elections in 1800

Turnout was calculated as the ratio between the total number of votes cast and population of free white males over the age of sixteen.

```{r fig.width=8, fig.height=8}
color_turnout <- colorBin("Blues", domain = counties_1800$turnout,
                              na.color = "darkgrey")

leaflet(counties_1800) %>% 
  addPolygons(stroke = TRUE, 
              color = "#000000", 
              opacity = 1, 
              weight = 1,
              smoothFactor = 0.5,
              fillColor = ~color_turnout(turnout),
              fillOpacity = 1, 
              popup = ~paste0("<b>County: </b>", county, "<br>",
                              "Federalist vote: ",
                              prettyNum(Federalist, big.mark = ","),
                              " (", round(federalist_percentage * 100, 1),
                              "%)<br>", "Republican vote: ",
                              prettyNum(Republican, big.mark = ","),
                              " (", round((1 - federalist_percentage) * 100, 1),
                              "%)<br>",
                              "Turnout: ", round(turnout * 100, 1), "%")) %>%
  addPolygons(data = districts_1800,
              stroke = TRUE,
              color = "#000000",
              opacity = 1, 
              fill = FALSE, 
              weight = 3,
              smoothFactor = 0.5) %>%
  addLegend(pal = color_turnout,
            values = ~turnout,
            labFormat = labelFormat(suffix = "%",
                                    transform = function(x) 100 * x),
            na.label = "NA",
            title = "Turnout")
```

### Sources

Election records: Philip Lampi Collection of American Electoral Returns, 1788--1825, American Antiquarian Society (2007). And *A New Nation Votes: American Election Returns, 1787--1825*: <http://elections.lib.tufts.edu/>.

Population data: Minnesota Population Center, *National Historical Geographic Information System*, version 2.0, University of Minnesota (2011): <https://nhgis.org/>.

County boundary data: John H. Long, Peggy Tuck Sinko, et al. *Atlas of Historic County Boundaries*, Dr. William M. Scholl Center for American History and Culture, Newberry Library (2010): <http://publications.newberry.org/ahcbp/project.html>.

Congressional district boundary data: Jeffrey B. Lewis, Brandon DeVine, and Lincoln Pritcher with Kenneth C. Martis, "United States Congressional District Shapefiles," University of California, Los Angeles (2013): <http://cdmaps.polisci.ucla.edu/>.




